---
title: "Pre-election data EDA"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
---

# Exploratory data analysis plotting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
knitr::opts_chunk$set(fig.width = 12, fig.height = 8)
options(knitr.table.format = "html")
library(tidyverse)
library(viridis)
library(here)
library(ggthemes)
library(knitr)
library(jsonlite)
library(patchwork)

theme_set(theme_bw())

response_type_colors <- c(
  "Harris" = "darkblue", "pronoun female" = "blue",
  "Trump" = "darkred", "pronoun male" = "red",
  "hedged" = "darkgreen", "pronoun neutral" = "green",
  "ambiguous NP" = "lightgreen",
  "OTHER" = "gray"
)
scale_fill_response_types <- scale_fill_manual(
  values = response_type_colors,
  breaks = names(response_type_colors)
)
candidate_colors <- c(
  "Harris" = "darkblue", "Trump" = "darkred",
  "other" = "gray"
)
pronoun_colors <- c(
  "she" = "blue", "he" = "red",
  "they" = "green"
)
scale_color_pronouns <- scale_color_manual(
  values = pronoun_colors,
  breaks = names(pronoun_colors)
)

recall_question <- read_csv(here("data/processed/pre_recall_question.csv"))
expectations <- read_csv(here("data/processed/pre_expectations.csv")) |>
  mutate(
    candidate = recode(candidate, harris = "Harris", trump = "Trump"),
    candidate = factor(candidate, levels = c("Harris", "Trump", "other"))
  )
cloze <- read_csv(here("data/processed/pre_cloze.csv")) |> mutate(
  response_type = factor(response_type,
    levels = c(
      "Harris", "pronoun female", "Trump", "pronoun male",
      "hedged", "pronoun neutral", "ambiguous NP", "OTHER"
    )
  )
)
spr <- read_csv(here("data/processed/pre_spr.csv")) |> # RTs are residualized
  mutate(
    nth_pronoun = factor(case_when(is_pron1 ~ 1, is_pron2 ~ 2, .default = NA)),
    pronoun_gender = factor(case_when(is_pron1 ~ type1, is_pron2 ~ type2, .default = NA))
  )
# spr_all <- read_csv(here("data/processed/pre_spr_all.csv"))
maze <- read_csv(here("data/processed/pre_maze.csv")) |> # Maze RTs are not residualized
  mutate(
    nth_pronoun = factor(case_when(is_pron1 ~ 1, is_pron2 ~ 2, .default = NA)),
    pronoun_gender = factor(case_when(is_pron1 ~ type1, is_pron2 ~ type2, .default = NA))
  )
participants <- read_csv(here("data/processed/pre_participants.csv"))
```

```{r}
np_cloze <- n_distinct(cloze$workerid)
np_maze <- n_distinct(maze$workerid)
np_spr <- n_distinct(spr$workerid)
message("Number of cloze participants: ", np_cloze)
message("Number of maze participants : ", np_maze)
message("Number of SPR participants  : ", np_spr)
message("                       total: ", np_cloze + np_maze + np_spr)
```

## Expectations

Check that we got a rough balance in event probability responses for the two candidates (given the uncertainty in polling and prediction markets, we should expect respondents to reflect this).
```{r fig.height=4}
# expectations |>
#   ggplot(aes(response,topic)) +
#   labs(y='candidate',title="responded probability of being next president") +
#   geom_boxplot(draw_quantiles=c(.5))

expectations |> mutate(order=if_else(str_starts(condition, "event"), "event-trial", "trial-event")) |>
  ggplot(aes(probability, candidate, color = candidate)) +
  labs(
    y = NULL,
    title = "\"Who do you think will be the US president in February 2025?\"",
    subtitle = "responses normalized per participant"
  ) +
  scale_color_manual(values = candidate_colors) +
  geom_violin(draw_quantiles = .5, alpha=0.5) +
  facet_grid(order~.) +
  geom_jitter(size = 0.5, alpha = 0.5, color = "gray") +
  stat_summary(fun.data = "mean_se", geom = "pointrange")
```

Looking for order effects, (combining all orders that start with "event" to one, for this plot, since all are identical wrt this data)
```{r fig.height=6}
expectations |>
  mutate(condition = factor(case_when(
    condition |> str_starts("event-") ~ "first",
    .default = paste("after", str_replace(condition, "-event", ""))
  ))|> fct_relevel("first")) |>
  ggplot(aes(probability, candidate, color = candidate)) +
  labs(
    y = NULL,
    title = "\"Who do you think will be the US president in February 2025?\"",
    subtitle = "responses normalized per participant, faceted by order"
  ) +
  scale_color_manual(values = candidate_colors) +
  facet_grid(condition ~ .) +
  # geom_jitter(size = 0.5, alpha = 0.5, color = "gray") +
  stat_summary(fun.data = "mean_se", geom = "pointrange")
```

## Cloze

```{r fig.height=6}
cloze_categorized <- cloze |> # filter(response_type!="OTHER") |>
  mutate(category = case_when(
    response_type %in% c("Harris", "pronoun female") ~ "female",
    response_type %in% c("Trump", "pronoun male") ~ "male",
    response_type %in% c("hedged", "pronoun neutral", "ambiguous NP") ~ "neutral",
    .default = "OTHER"
  ))
cloze_categorized |>
  ggplot(aes(category, fill = response_type)) +
  geom_bar() + # facet_wrap(~item) +
  geom_text(stat = "count", aes(label = ..count..), color = "white", position = position_stack(vjust = 0.5)) +
  scale_fill_response_types +
  labs(title = "Type of cloze responses") +
  theme(legend.position = "top", legend.direction = "horizontal")
```

Faceting by order:
```{r fig.height=6}
cloze_categorized |>
  ggplot(aes(category, fill = response_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = ..count..), color = "white", position = position_stack(vjust = 0.5)) +
  facet_wrap(~condition, scales = "free_y") +
  scale_fill_response_types +
  labs(title = "Type of cloze responses, faceted by order") +
  theme(legend.position = "top", legend.direction = "horizontal")
```
In the event-cloze order there seems to be a somewhat larger bias toward producing male pronouns or male referents/Trump.

_But it seems the difference between orders here is actually not significant, per Titus' binomial 95%-HPDI test:_
```{r fig.width = 5, fig.height = 5, out.width="50%"}
# remotes::install_github("tmalsburg/binomialCRIs")
if (require("binomialCRIs")) {
  F_pre <- nrow(filter(cloze, condition=="cloze-event", response_type=="pronoun female"))
  M_pre <- nrow(filter(cloze, condition=="cloze-event", response_type=="pronoun male"))
  F_post <- nrow(filter(cloze, condition=="event-cloze", response_type=="pronoun female"))
  M_post <- nrow(filter(cloze, condition=="event-cloze", response_type=="pronoun male"))
  plot_binomial_hpdi(
    n_successes = F_pre, n_trials = F_pre + M_pre,
    prob = 0.95
  ) +
    geom_vline(xintercept = F_post / (F_post + M_post)) +
    labs(
      title = "Binomial HDPI check",
      subtitle = paste(
        "posterior for proportion fem. in cloze-event\n",
        "vertical line at proportion fem. in event-cloze"
      )
    )
}
```

Breaking down further by gender:
```{r}
cloze |>
  left_join(participants) |> # filter(response_type!="OTHER") |>
  mutate(category = case_when(
    response_type %in% c("Harris", "pronoun female") ~ "female",
    response_type %in% c("Trump", "pronoun male") ~ "male",
    response_type %in% c("hedged", "pronoun neutral", "ambiguous NP") ~ "neutral",
    .default = "OTHER"
  )) |>
  filter(gender %in% c("Male", "Female")) |>
  ggplot(aes(category, fill = response_type)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = ..count..), color = "white", position = position_stack(vjust = 0.5)) +
  facet_wrap(gender ~ condition, scales = "free_y") +
  scale_fill_response_types +
  labs(title = "Type of cloze responses, faceted by order and gender") +
  theme(legend.position = "top", legend.direction = "horizontal")
```
^ It appears the effect of order on bias may be driven by gender: 
Male-identifying participants produced more male-referring cloze completions 
when the cloze completion task followed event-estimation, compared to when cloze completion came first.

For female-identifying participants, there is no such marked difference between the orders.

Across participant-genders, there is a smaller proportion of neutral/hedged referents when completing sentences after 
event estimation, compared to before event-estimation.

_But this is all something that would take larger sample size to actually assess._

```{r}
cloze |> # filter(response_type!="OTHER") |>
  mutate(category = case_when(
    response_type %in% c("Harris", "pronoun female") ~ "female",
    response_type %in% c("Trump", "pronoun male") ~ "male",
    response_type %in% c("hedged", "pronoun neutral", "ambiguous NP") ~ "neutral",
    .default = "OTHER"
  )) |>
  mutate(partial = gsub("The next US president will be sworn into office in January 2025. ", "", partial)) |>
  ggplot(aes(category, fill = response_type)) +
  geom_bar() +
  facet_wrap(ncol = 4, ~ paste0(
    sprintf("%02d", item), ":\n",
    gsub("(.{40})", "\\1\n", partial)
  )) +
  scale_fill_response_types +
  labs(title = "Type of cloze responses, by item") +
  theme(legend.position = "top", legend.direction = "horizontal")
```

## Maze
Look at average Maze RT by gender of pronoun, and for each condition of two sentences presented.
Shows an effect of gender (male pronoun is read faster than female, neutral is perhaps even faster than male).
These Maze RTs are not residualized. But could be for slightly smaller std error probably.
```{r}
# maze |>
#   count(item1, item2, type, type1, type2) |>
#   ggplot(aes(factor(item1), factor(item2), size = n, color = type)) +
#   geom_point() +
#   facet_grid(type1 ~ type2)
```

```{r}
p_maze1 <- maze |>
  filter(correct, is_target) |>
  ggplot(aes(pronoun_gender, log(rt), color = factor(pronoun_gender))) +
  stat_summary() +
  scale_color_pronouns +
  labs(title = "Maze RT on pronouns by gender", color = "pronoun gender", subtitle = "overall")

p_maze2 <- maze |>
  filter(is_target) |>
  filter(correct) |>
  ggplot(aes(nth_pronoun, log(rt), color = factor(pronoun_gender))) +
  stat_summary(position = position_dodge(width = .2)) +
  facet_grid(~type, scales = "free") +
  scale_color_pronouns +
  labs(
    title = "Maze RT on pronouns by stimulus type", color = "pronoun gender",
    subtitle = "(gender of pronoun in each of two sentences read back to back)"
  )

p_maze <- (p_maze1 / p_maze2) + plot_layout(guides = "collect")
p_maze
```

Faceting by order
```{r fig.height=4}
maze |>
  filter(correct, is_target) |>
  ggplot(aes(pronoun_gender, log(rt), color = factor(pronoun_gender))) +
  stat_summary() +
  scale_color_pronouns +
  facet_grid(~condition, scales = "free") +
  labs(title = "Maze RT on pronouns by gender", subtitle = "facet by order", color = "pronoun gender")
```

There is an overall less pronounced bias for male pronoun for participants who 
did the maze task following the event estimation (event-maze) versus before (maze-event).

Faceting also by participant gender
```{r fig.height=6}
participants |>
  select(workerid, gender) |>
  inner_join(maze, join_by(workerid)) |>
  filter(correct, is_target, gender %in% c("Female", "Male")) |>
  ggplot(aes(pronoun_gender, log(rt), color = factor(pronoun_gender))) +
  stat_summary() +
  facet_grid(condition ~ gender) +
  scale_color_pronouns +
  labs(
    title = "Maze RT on pronouns by participant's reported gender",
    color = "pronoun gender", subtitle = "overall"
  )
```
^ Breaking down by gender, it seems the order-dependence is mostly driven by males.  
For females, the processing advantage for male pronoun over female is similar whether event-estimation was done before or after the reading task, 
whereas for males, the bias is much larger in maze-event versus event-maze.


And broken down by participant reported political affiliation
```{r fig.height=6}
participants |>
  select(workerid, political_aff) |>
  inner_join(maze, join_by(workerid)) |>
  filter(correct, is_target, political_aff %in% c("Democrat", "Republican", "Independent")) |>
  ggplot(aes(pronoun_gender, log(rt), color = factor(pronoun_gender))) +
  stat_summary() +
  facet_grid(condition ~ political_aff) +
  scale_color_pronouns +
  labs(
    title = "Maze RT on pronouns by participant's reported political affil.",
    color = "pronoun gender", subtitle = "overall"
  )
```
^ Breaking down by political affiliation we see something similar, it seems the order-dependence is mostly non-democrats. 
For democrats, the processing advantage for male pronoun over female is similar whether event-estimation was done before or after the reading task, 
whereas for independents and Republicans, the bias is pronounced only when the maze task came before event estimation.

And broken down by participant reported preference
```{r fig.height=6}
participants |>
  select(workerid, prefer) |>
  inner_join(maze, join_by(workerid)) |>
  filter(correct, is_target) |>
  mutate(prefer = if_else(!(prefer %in% c("Donald Trump", "Kamala Harris")), "Refuse/No pref./NA", prefer)) |>
  ggplot(aes(pronoun_gender, log(rt), color = factor(pronoun_gender))) +
  stat_summary() +
  facet_grid(condition ~ prefer) +
  scale_color_pronouns +
  labs(
    title = "SPR RT on pronouns by participant's reported candidate preference",
    color = "pronoun gender", subtitle = "overall"
  )
```
^ Interestingly (?) the order-effect is not as clearly explainable by candidate preference.

We can also look at the effect of event probability (who the participant thinks will be president) on pronoun RT bias
```{r}
# Plot Maze RT on pronouns by gender against belief that president will be female
expectations |>
  pivot_wider(names_from = candidate, values_from = c(probability, response)) |>
  select(workerid, starts_with("probability")) |>
  left_join(maze) |>
  filter(correct, is_target) |>
  ggplot(aes(probability_Harris, log(rt), color = pronoun_gender)) +
  geom_smooth(method = "lm") +
  scale_color_pronouns +
  labs(
    title = "Maze RT on pronouns, plotted against probability of female president",
    x = "belief that president will be Harris", y = "Maze RT (log scale)"
  )
```


Or, we could look at participant age
```{r}
participants |>
  select(workerid, gender, age) |>
  inner_join(maze, join_by(workerid)) |>
  filter(correct, is_target, age < 100) |>
  ggplot(aes(age, log(rt), color = pronoun_gender)) +
  geom_smooth(method = "lm") +
  # geom_point(alpha=0.25) +
  scale_color_pronouns +
  # facet_grid(condition ~ .) +
  labs(
    title = "Maze RT on pronouns, plotted against participant age",
    x = "age", y = "Maze RT (log scale)"
  )
```

----

```{r fig.height=12}
maze |>
  mutate(
    full_type = type, full_item = item,
    s = case_when(in_s == 0 ~ s0, in_s == 1 ~ s1, in_s == 2 ~ s2),
    type = case_when(in_s == 0 ~ NA, in_s == 1 ~ type1, in_s == 2 ~ type2),
    item = case_when(in_s == 0 ~ 0, in_s == 1 ~ item1, in_s == 2 ~ item2),
  ) |>
  filter(
    correct,
    # Filter to not get effects from switching pronouns midstream
    # (only use first sentence)
    in_s == 1
  ) |>
  select(-c(type1, type2, item1, item2)) |>
  summarize(
    # summarize by `in_s` or `full_type` to get sentence order effect
    .by = c(item, type, word_number_in_s, is_target, word),
    n = n(),
    m = mean(log(rt)),
    se = sd(log(rt)) / sqrt(n)
  ) |>
  filter(item != 0) |>
  ggplot(aes(
    word_number_in_s,
    m, ymin = m - 1.96 * se, ymax = m + 1.96 * se,
    group = type, color = type
  )) +
  geom_line(alpha = 0.5) +
  scale_color_pronouns + 
  geom_pointrange(aes(size = is_target, shape = is_target), alpha = 0.5) +
  scale_size_manual(values = c(0.5, 1)) +
  geom_text(aes(label=word,y=-Inf), alpha=0.33, angle = 15, vjust = -1, hjust = 0.5, size = 2) +
  facet_grid(item ~ .) +
  labs(title="Maze", y="log RT")
```

## SPR

Likewise look at average SPR RT by gender of pronoun, and for each condition of two sentences presented.
These RTs are residualized.
Results look similar to Maze RTs, but are less pronounced/lower power.

Still, I think high enough power that we don't need to run further trials.

```{r}
# spr |>
#   count(item1, item2, type, type1, type2) |>
#   ggplot(aes(factor(item1), factor(item2), size = n, color = type)) +
#   geom_point() +
#   facet_grid(type1 ~ type2)

p_spr1 <- spr |>
  filter(is_target) |>
  ggplot(aes(pronoun_gender, log(rt_gmean), color = factor(pronoun_gender))) +
  stat_summary() +
  scale_color_pronouns +
  labs(
    title = "SPR RT on pronouns by gender", color = "pronoun gender",
    subtitle = "overall (using RTs averaged across 4-word window for spillover)",
    y = "log residualized RT"
  )


p_spr2 <- spr |>
  filter(is_target) |>
  ggplot(aes(nth_pronoun, log(rt_gmean), color = factor(pronoun_gender))) +
  stat_summary(position = position_dodge(width = .2)) +
  facet_grid(~type, scales = "free") +
  scale_color_pronouns +
  labs(
    title = "SPR RT on pronouns by  stimulus type", color = "pronoun gender",
    subtitle = "(RTs averaged across 4-word window for spillover)\n(gender of pronoun in each of two sentences read back to back)",
    y = "log residualized RT"
  )
p_spr <- (p_spr1 / p_spr2) + plot_layout(guides = "collect")
p_spr
```

```{r fig.height=4}
spr |>
  filter(is_target) |>
  ggplot(aes(pronoun_gender, log(rt_gmean), color = factor(pronoun_gender))) +
  stat_summary() +
  scale_color_pronouns +
  facet_grid(~condition, scales = "free") +
  labs(
    title = "SPR RT on pronouns by gender", color = "pronoun gender",
    subtitle = "overall (using RTs averaged across 4-word window for spillover)",
    y = "log residualized RT"
  )
```

As with the Maze task, the bias toward male pronouns is less when the reading task follows event-probability-estimation (event-spr order) versus when it precedes it (spr-event).
In this case, there seems to be no bias at all for male vs female pronoun in the event-spr order.

```{r}
## Inspect:
## Comparing different lags,
## rather than just using the mean over the whole window
rt_lag_colors <- c(
  `w_{i}` = "black",
  `w_{i+1}` = "darkgray",
  `w_{i+2}` = "gray",
  `w_{i+3}` = "lightgray",
  avg = "purple"
)
scale_color_rt_lag <- scale_color_manual(
  values = rt_lag_colors,
  breaks = names(rt_lag_colors)
)

spr_longer <- spr |>
  select(-rt.raw) |>
  filter(is_target) |>
  pivot_longer(
    cols = starts_with("rt"), names_to = "RT lag", values_to = "reading_time"
  ) |>
  mutate(`RT lag` = case_when(
    `RT lag` == "rt_gmean" ~ "avg",
    str_starts(`RT lag`, "rt_") ~ paste(gsub("rt_(.)next", "w_{i+\\1}", `RT lag`)),
    `RT lag` == "rt" ~ "w_{i}",
  ))

p_spr_lags1 <- spr_longer |>
  ggplot(aes(`RT lag`, log(reading_time), color = `RT lag`)) +
  stat_summary() +
  facet_grid(condition ~ pronoun_gender, scales = "free") +
  scale_color_rt_lag +
  labs(
    title = "SPR by gender",
    subtitle = "Looking at RT with different lags w_{i+n}, where w_i = target pronoun. \nAlso showing geometric_mean(w_{i+n}) for n = {0,...,3}",
    y = "residualized log RT"
  )

p_spr_lags2 <- spr_longer |>
  summarize(
    .by = c(`RT lag`, nth_pronoun, type, condition),
    m = mean(log(reading_time)),
    se = sd(log(reading_time)) / sqrt(n())
  ) |>
  ggplot(aes(x = nth_pronoun, group = `RT lag`, y = m, ymin = m - se, ymax = m + se, color = `RT lag`)) +
  geom_pointrange(position = position_dodge(width = .4)) +
  facet_grid(condition ~ type, scales = "free") +
  geom_line() +
  scale_color_rt_lag +
  labs(
    title = "Maze RT on pronouns by stimulus type",
    subtitle = "(gender of pronoun in each of two sentences read back to back)",
    y = "residualized log RT"
  )
p_spr_lags <- (p_spr_lags1 / p_spr_lags2) + plot_layout(guides = "collect")
p_spr_lags
```


<!-- We could also try to look at the effect of event belief, but this is too low power with SPR:  -->
```{r}
# # Plot Maze RT on pronouns by gender against belief that president will be female
# expectations |>
#   pivot_wider(names_from = candidate, values_from = c(probability, response)) |>
#   select(workerid, starts_with("probability")) |>
#   # left_join(demographics_and_prolific_uniq) |>
#   left_join(spr) |>
#   filter(is_target) |>
#   ggplot(aes(probability_Harris, log(rt_gmean), color = pronoun_gender)) +
#   geom_smooth(method = "lm") +
#   # geom_point(alpha=0.25) +
#   labs(
#     title = "SPR RT on pronouns, plotted against probability of female president",
#     x = "belief that president will be Harris", y = "Maze RT"
#   )
```




```{r fig.height=4}
participants |>
  select(workerid, gender) |>
  inner_join(spr, join_by(workerid)) |>
  filter(is_target, gender %in% c("Female", "Male")) |>
  ggplot(aes(pronoun_gender, log(rt_gmean), color = factor(pronoun_gender))) +
  stat_summary() +
  facet_grid(~gender) +
  scale_color_pronouns +
  labs(
    title = "SPR RT on pronouns by participant's reported gender",
    color = "pronoun gender", subtitle = "overall"
  )
```

^ SPR averages broken down by participant gender

```{r fig.height=6}
participants |>
  select(workerid, political_aff) |>
  inner_join(spr, join_by(workerid)) |>
  filter(is_target, political_aff %in% c("Democrat", "Republican", "Independent")) |>
  ggplot(aes(pronoun_gender, log(rt_gmean), color = factor(pronoun_gender))) +
  stat_summary() +
  facet_grid(~political_aff) +
  scale_color_pronouns +
  labs(
    title = "SPR RT on pronouns by participant's reported political affil.",
    color = "pronoun gender", subtitle = "overall"
  )
```
^ SPR averages broken down by participant political affiliation
----

```{r fig.height=12}
spr |>
  mutate(
    full_type = type, full_item = item,
    s = case_when(in_s == 0 ~ s0, in_s == 1 ~ s1, in_s == 2 ~ s2),
    type = case_when(in_s == 0 ~ NA, in_s == 1 ~ type1, in_s == 2 ~ type2),
    item = case_when(in_s == 0 ~ 0, in_s == 1 ~ item1, in_s == 2 ~ item2),
  ) |>
  filter(
    # Filter to not get effects from switching pronouns midstream
    type1==type2
  ) |>
  select(-c(type1, type2, item1, item2)) |>
  summarize(
    # summarize by `in_s` or `full_type` to get sentence order effect
    .by = c(item, type, word_number_in_s, is_target, word),
    n = n(),
    m = mean(log(rt)),
    se = sd(log(rt)) / sqrt(n)
  ) |>
  filter(item != 0) |>
  ggplot(aes(
    word_number_in_s,
    m, ymin = m - 1.96 * se, ymax = m + 1.96 * se,
    group = type, color = type
  )) +
  geom_line(alpha = 0.5) +
  scale_color_pronouns + 
  geom_pointrange(aes(size = is_target, shape = is_target), alpha = 0.5) +
  scale_size_manual(values = c(0.5, 1)) +
  geom_text(aes(label=word,y=-Inf), alpha=0.33, angle = 15, vjust = -1, hjust = 0.5, size = 2) +
  facet_grid(item ~ .) +
  labs(title="SPR", y="log RT")
```